A Gaussian process model of human electrocorticographic data

### 2020-02-11 paper/main.tex
minor updates to paper and correlation between best locations by frequency

\section*{Approach}
The SuperEEG approach to inferring high temporal resolution full-brain
activity patterns is outlined and summarized in
Figure~\ref{fig:methods}. We describe (in this section) and evaluate
(in \textit{Results}) our approach using a two large previously
collected dataset comprising multi-session intracranial recordings.
Dataset 1 comprises multi-session recordings taken from 6876
electrodes implanted in the brains of 88 epilepsy
patients~\citep{SedeEtal03, SedeEtal07a, SedeEtal07b, MannEtal11,
  MannEtal12}.  Each recording session lasted from 0.2--3~h (total
recording time: 0.3--14.2~h; Fig.~\suppstats E).  During each
recording session, the patients participated in a free recall list
learning task, which lasted for up to approximately 1~h.  In addition,
the recordings included ``buffer'' time (the length varied by patient)
before and after each experimental session, during which the patients
went about their regular hospital activities (confined to their
hospital room, and primarily in bed).  These additional activities
included interactions with medical staff and family, watching
television, reading, and other similar activities.  For the purposes
of the Dataset 1 analyses presented here, we aggregated all data
across each recording session, including recordings taken during the
main experimental task as well as during non-experimental time.  We
used Dataset 1 to develop our main SuperEEG approach, and to examine
the extent to which SuperEEG might be able to generate task-general
predictions.  Dataset 2 comprised multi-session recordings from 4436
electrodes implanted in the brains of 40 epilepsy
patients~\citep{EzzyEtal17, HoraEtal17, KragEtal17, KuceEtal17,
  LinEtal17, SoloEtal18, WeidEtal18, EzzyEtal18, KuceEtal18}.  Each
recording session lasted from 0.4--2.2~h (total recording time:
0.4--6.6~h; Fig.~\suppstats K).  Whereas Dataset 1 included recordings
taken as the patients participated in a variety of activities, Dataset
2 included recordings taken as each patient performed each of two
specific experimental memory tasks: a random word list free recall
task (Experiment 1) and a categorized word list free recall task
(Experiment 2).  We used Dataset 2
to further examine the ability of SuperEEG to generalize its
predictions within versus across tasks.  Figure~\suppstats~provides
additional information about both datasets.

@@

We first applied fourth order Butterworth notch filter to remove 60 Hz
($\pm$ .5 Hz) line noise from every recording (from every electrode).
Next, we downsampled the recordings (regardless of the original
samplerate) to 250 Hz.  (This downsampling step served to both
normalize for differences in sampling rates across patients and to
ease the computational burden of our subsequent analyses.)  We then
excluded any electrodes that showed putative epileptiform activity.
Specifically, we excluded from further analysis any electrode that
exhibited an maximum kurtosis of 10 or greater across all of that
patient's recording sessions.  We also excluded any patients with
fewer than 2 electrodes that passed this criteria, as the SuperEEG
algorithm requires measuring correlations between 2 or more electrodes
from each patient.  For Dataset 1, this yielded clean recordings from
4168 electrodes implanted throughout the brains of 67 patients
(Fig.~\ref{fig:methods}A); for Dataset 2, this yielded clean
recordings from 5023 electrodes from 78 patients.  Each individual
patient contributes electrodes from a limited set of brain locations,
which we localized in a common space~\citep[MNI152;][]{GrabEtal06}; an
example Dataset 1 patient's 54 electrodes that passed the above
kurtosis threshold test are highlighted in black and red.

@@

We used this general approach to
quantify the algorithm's performance across the full dataset.For each held-out electrode, from each held-out patient in turn, we
computed the average correlation (across recording sessions) between
the SuperEEG-reconstructed voltage traces and the observed voltage
traces from that electrode.  For this analysis we set $\bar{R}$ to be
the union of all electrode locations across all patients.  This
yielded a single correlation coefficient for each electrode location
in $\bar{R}$, reflecting how well the SuperEEG algorithm was able to
recover the recording at that location by incorporating data across
patients (black histogram in Fig.~\ref{fig:corrmap}A, map in
Fig.~\ref{fig:corrmap}C).  The observed distribution of correlations
was centered well above zero (mean: 0.51; $t$-test comparing mean of
distribution of $z$-transformed average patient correlation coefficients
to 0: $t(66) = 23.55, p < 10^{-10}$), indicating that the SuperEEG algorithm
recovers held-out activity patterns substantially better than random
guessing.

@@

We then wondered which
networks were involved in each of
these frequency-specific intersecting regions.  To better understand the networks involved
we colored the regions (Fig.~\ref{fig:networks}A) based on the
network parcellation ~\citep[using $k$=7;][]{YeoEtal11}} (Fig.~\ref{fig:networks}D) and calculated the proportion of each
of those networks (Fig.~\ref{fig:networks}B).  Consistent with the idea outlined
above, we found networks most involved in the lowest frequency band,
$\delta$, were networks involved integration and global processing
(specifically the dorsal attention ~\citep{MarzEtal13}, default mode,
and frontoparietal networks).  Although these networks are also seen
in the $\theta$ intersection, they are less proportionally to the
visual network.  Alpha power is consistently modulated in EEG studies
of attention and visuo-motor ~\citep{MantEtal07, MarzEtal13}, and we
similarly found a high proportion of visual, somatomotor, and dorsal attentio networks
within this band.  We found $\beta$ intersections to predominantly
consist of somatomotor network, in line with~\citep{RittEtal09, VaneEtal11}.  Also accordant with heavily cited
literature~\citep{FrieEtal07}, the network most predominant in the
$\gamma$ intersection was the visual network.  Lastly, broadband power
was found predominantly in frontoparietal and default mode networks,
consistent with ~\citep{OssaEtal11}.

>>>>

\section*{Approach}
The SuperEEG approach to inferring high temporal resolution full-brain
activity patterns is outlined and summarized in
Figure~\ref{fig:methods}. We describe (in this section) and evaluate
(in \textit{Results}) our approach using a two large previously
collected datasets comprising multi-session intracranial recordings.
Dataset 1 comprises multi-session recordings taken from 6876
electrodes implanted in the brains of 88 epilepsy
patients~\citep{SedeEtal03, SedeEtal07a, SedeEtal07b, MannEtal11,
  MannEtal12}.  Each recording session lasted from 0.2--3~h (total
recording time: 0.3--14.2~h; Fig.~\suppstats E).  During each
recording session, the patients participated in a free recall list
learning task, which lasted for up to approximately 1~h.  In addition,
the recordings included ``buffer'' time (the length varied by patient)
before and after each experimental session, during which the patients
went about their regular hospital activities (confined to their
hospital room, and primarily in bed).  These additional activities
included interactions with medical staff and family, watching
television, reading, and other similar activities.  For the purposes
of the Dataset 1 analyses presented here, we aggregated all data
across each recording session, including recordings taken during the
main experimental task as well as during non-experimental time.  We
used Dataset 1 to develop our main SuperEEG approach, and to examine
the extent to which SuperEEG might be able to generate task-general
predictions.  Dataset 2 comprised multi-session recordings from 4436
electrodes implanted in the brains of 40 epilepsy
patients~\citep{EzzyEtal17, HoraEtal17, KragEtal17, KuceEtal17,
  LinEtal17, SoloEtal18, WeidEtal18, EzzyEtal18, KuceEtal18}.  Each
recording session lasted from 0.4--2.2~h (total recording time:
0.4--6.6~h; Fig.~\suppstats K).  Whereas Dataset 1 included recordings
taken as the patients participated in a variety of activities, Dataset
2 included recordings taken as each patient performed each of two
specific experimental memory tasks: a random word list free recall
task (Experiment 1) and a categorized word list free recall task
(Experiment 2).  We used Dataset 2
to further examine the ability of SuperEEG to generalize its
predictions within versus across tasks.  Figure~\suppstats~provides
additional information about both datasets.

@@

We first applied fourth order Butterworth notch filter to remove 60 Hz
($\pm$ .5 Hz) line noise from every recording (from every electrode).
Next, we downsampled the recordings (regardless of the original
sample rate) to 250 Hz.  (This downsampling step served to both
normalize for differences in sampling rates across patients and to
ease the computational burden of our subsequent analyses.)  We then
excluded any electrodes that showed putative epileptiform activity.
Specifically, we excluded from further analysis any electrode that
exhibited an maximum kurtosis of 10 or greater across all of that
patient's recording sessions.  We also excluded any patients with
fewer than 2 electrodes that passed this criteria, as the SuperEEG
algorithm requires measuring correlations between 2 or more electrodes
from each patient.  For Dataset 1, this yielded clean recordings from
4168 electrodes implanted throughout the brains of 67 patients
(Fig.~\ref{fig:methods}A); for Dataset 2, this yielded clean
recordings from 5023 electrodes from 78 patients.  Each individual
patient contributes electrodes from a limited set of brain locations,
which we localized in a common space~\citep[MNI152;][]{GrabEtal06}; an
example Dataset 1 patient's 54 electrodes that passed the above
kurtosis threshold test are highlighted in black and red.

@@

We used this general approach to
quantify the algorithm's performance across the full dataset. For each held-out electrode, from each held-out patient in turn, we
computed the average correlation (across recording sessions) between
the SuperEEG-reconstructed voltage traces and the observed voltage
traces from that electrode.  For this analysis we set $\bar{R}$ to be
the union of all electrode locations across all patients.  This
yielded a single correlation coefficient for each electrode location
in $\bar{R}$, reflecting how well the SuperEEG algorithm was able to
recover the recording at that location by incorporating data across
patients (black histogram in Fig.~\ref{fig:corrmap}A, map in
Fig.~\ref{fig:corrmap}C).  The observed distribution of correlations
was centered well above zero (mean: 0.51; $t$-test comparing mean of
distribution of $z$-transformed average patient correlation coefficients
to 0: $t(66) = 23.55, p < 10^{-10}$), indicating that the SuperEEG algorithm
recovers held-out activity patterns substantially better than random
guessing.

@@

We then wondered which
networks were involved in each of
these frequency-specific intersecting regions.  To better understand the networks involved
we colored the regions (Fig.~\ref{fig:networks}A) based on the
network parcellation ~\citep[using $k$=7;][]{YeoEtal11}} (Fig.~\ref{fig:networks}D) and calculated the proportion of each
of those networks (Fig.~\ref{fig:networks}B).  Consistent with the idea outlined
above, we found networks most involved in the lowest frequency band,
$\delta$, were networks involved integration and global processing
(specifically the dorsal attention ~\citep{MarzEtal13}, default mode,
and frontoparietal networks).  Although these networks are also seen
in the $\theta$ intersection, they are less proportionally to the
visual network.  Alpha power is consistently modulated in EEG studies
of attention and visuo-motor ~\citep{MantEtal07, MarzEtal13}, and we
similarly found a high proportion of visual, somatomotor, and dorsal attention networks
within this band.  We found $\beta$ intersections to predominantly
consist of somatomotor network, in line with~\citep{RittEtal09, VaneEtal11}.  Also accordant with heavily cited
literature~\citep{FrieEtal07}, the network most predominant in the
$\gamma$ intersection was the visual network.  Lastly, broadband power
was found predominantly in frontoparietal and default mode networks,
consistent with ~\citep{OssaEtal11}.
